|

### 2023-05-12 paper/00-main.Rmd
more

abstract: |
  In American football, a pass rush is an attempt by the defensive team to disrupt the offense and prevent the quarterback (QB) from completing a pass. Existing metrics for assessing pass rush are either discrete-time quantities or based on subjective judgement. Using player tracking data, we propose STRAIN, a novel metric for evaluating pass rushers in the National Football League (NFL) at the continuous-time within-play level. Inspired by the concept of strain rate in materials science, STRAIN is a simple and interpretable means for measuring defensive pressure in football. It is a directly-observed statistic as a function of two features: the distance between the pass rusher and QB, and the rate at which this distance is being reduced. Our metric possesses great predictability of pressure and stability over time. We use a multilevel model to isolate the contribution of every pass rusher in a play to the average STRAIN while controlling for their pass blocker matchup, the QB, and play-level information. We apply our approach to NFL data and present results for the first 8 weeks of the 2021 season. In particular, we provide comparisons of STRAIN for different defensive positions and play outcomes, and rankings of the NFL's best pass rushers according to our metric.

@@

**GitHub repository:** https://github.com/getstrained/intro-strain.

**NFL Big Data Bowl 2023 Data:**: https://www.kaggle.com/competitions/nfl-big-data-bowl-2023/data.

>>>>

abstract: |
  In American football, a pass rush is an attempt by the defensive team to disrupt the offense and prevent the quarterback (QB) from completing a pass. Existing metrics for assessing pass rush performance are either discrete-time quantities or based on subjective judgement. Using player tracking data, we propose STRAIN, a novel metric for evaluating pass rushers in the National Football League (NFL) at the continuous-time within-play level. Inspired by the concept of strain rate in materials science, STRAIN is a simple and interpretable means for measuring defensive pressure in football. It is a directly-observed statistic as a function of two features: the distance between the pass rusher and QB, and the rate at which this distance is being reduced. Our metric possesses great predictability of pressure and stability over time. We use a multilevel model to isolate the contribution of every pass rusher in a play to the average STRAIN while controlling for their pass blocker matchup, the QB, and play-level information. We apply our approach to NFL data and present results for the first 8 weeks of the 2021 season. In particular, we provide comparisons of STRAIN for different defensive positions and play outcomes, and rankings of the NFL's best pass rushers according to our metric.

@@

All code related to this paper is available at https://github.com/getstrained/intro-strain. The data provided by the NFL Big Data Bowl 2023 is available at: https://www.kaggle.com/competitions/nfl-big-data-bowl-2023/data.

### 2023-05-12 paper/01-introduction.Rmd
more

In this work, using data made available in the 2023 BDB, we present a novel approach to measure the performance of pass rushers (defensive players whose role on a given play is to put pressure on the QB). Relative to other aspects of American football, such as QB evaluation [@burke2019deepqb, @reyers2021quarterback], the literature on evaluating pass rushers is scarce. Below, we provide a brief overview of existing pass rush metrics.

@@

Table \ref{tab:metrics} gives a summary of existing pass-rush metrics.
We now highlight what these quantities measure as well as their limitations.

Perhaps the most commonly-known statistics for evaluating defensive linemen on pass rush plays are sacks, hits, and hurries, which are observed at the discrete play-level.
Officially tracked by the NFL since 1982, a sack is recorded when a defensive player tackles the quarterback behind the line of scrimmage before the quarterback is able to throw a pass.
Sacks represent a simple binary measure of success for a rusher on a pass play.
However, there are many intermediate outcomes for a rusher that can also be considered successes that do not result in a sack.
Other traditional box-score statistics such hits and hurries are collected by various outlets where hits counts the number of times a QB is hit while passing and hurries counts how often a QB has to throw the ball earlier than they would have liked to.
The sum of sacks, hits, and hurries has been defined as "pressures" and is an improvement over sacks, but suffers from problems of subjectivity (e.g. was the play actually a hurry?) and are still binary outcomes (i.e. There is a continuum of plays that aren't hurries that all counted as simply "not hurries").
Pro football focus (PFF) comes up with a metric called "pass-rush productivity" (PRP), which is a minor modification from the aforementioned pressures metric (see Table \ref{tab:metrics}).
In particular, PRP gives twice as much weight to a sack relative to hurries and hits, which is a slight improvement to pressures.
However, the choice of weights is ad-hoc and still only measures binary outcomes similar to the shortcomings of previously discussed metrics.

@@

\begin{table}
\caption{A summary of previously-existing pass rush metrics. \label{tab:metrics}}
\centering
\begin{tabular}{p{0.27\textwidth}p{0.67\textwidth}}
\hline
Metric & Description \\ 
\hline
Sacks & A defender tackles the QB behind the line of scrimmage before the QB is able to throw a pass \\
Hits & A defender tackles the QB behind the line of scrimmage after the QB is able to throw a pass. \\
Hurries & A defender pressures the QB behind the line of scrimmage forcing the QB to throw the ball prior sooner than intended. \\
Pressures & $\text{Hurries} + \text{ Hits } + \text{ Sacks}$ \\
Pass-Rush Productivity & $\displaystyle \frac{(\text{Hurries} + \text{Hits})/2 + \text{Sacks}}{\text{ Pass Rush Snaps}}$\\
Time In Pocket & Time (in seconds) between ball snap and throw or pocket collapse for a QB \\
Pass Rush Win Rate & Rate at which pass rusher beats pass block within 2.5 seconds after ball snap\\
\hline
\end{tabular}
\end{table}

@@

The literature on the performance of linemen, either offensive or defensive in football is scant.
@AlamarGould2008 find an association between pass completion rate and successful pass blocking by offensive linemen.
The data for this study are collected for the first three weeks of the 2007 NFL season, manually recording whether a lineman holds a block and the time it took for quarterback to throw the football.
Later, @AlamarGoldner2011 use manually-tracked data during the 2010 season to estimate lineman performance for different team-positions instead of individual defenders (e.g. Chicago Bears' left tackle, Pittsburgh Steelers' center).
The paper uses survival analysis to model time in pocket and completion percentage for quarterbacks, before ultimately measures a lineman's contribution to the overall offense in terms of yards gained.

@@

<!-- [this next part needs re-wording] -->
<!-- They then go on to use a survival analysis model to estimate time in pocket.  The average probability of success from the survival model is then used as a covariate in a model to estimate the probability of competing a pass, and they show that there is a significant relationship between these two probabilities. -->

## Contributions of STRAIN

All the aforementioned methods suffered from not having the data that was actually necessary to develop a meaningful metric for linemen in football, either offensive or defensive.
Now though, thanks to the 2023 Big Data Bowl, we have a publicly available tracking data not only for the linemen, but for every player on the field.
This allows us to meaningfully assess the performance of linemen in football.
In this paper, we focus on defensive linemen, and propose STRAIN, a metric for pass rush evaluation inspired by the concept of strain rate in materials science.
STRAIN gives a continuous measure of pressure for each of the defensive linemen over the course of an entire play.
This allows for the measurement of pass rush success even on plays that do not result in an observed outcome like a sack, hit, or hurry.
We view this as a major step forward in the accurate evaluation of defensive linemen performance.
Further, we note than while we focus exclusively on defensive linemen here, the proposed metric STRAIN is relatively easily extended to the evaluation of offensive linemen performance.

The remainder of this paper is outlined as follows.
We first describe the player tracking data provided by 2023 Big Data Bowl in Section \ref{sec:data}.
We then introduces the mathematical motivation and definition of our metric STRAIN and our modeling approach in Section \ref{sec:methods}.
We present an real-game illustration of the proposed metric and study its performance in Section \ref{sec:results}.
We close with a discussion of future directions for STRAIN in Section \ref{sec:discussion}.

>>>>

In this work, using data made available in the 2023 BDB, we present a novel approach to measure the performance of pass rushers. Relative to other aspects of American football, such as QB evaluation [@burke2019deepqb, @reyers2021quarterback], the literature on evaluating pass rushers is scarce. Below, we provide a brief overview of existing pass rush metrics.

<!-- (defensive players whose role on a given play is to put pressure on the QB) -->

@@

Table \ref{tab:metrics} gives a summary of existing football metrics for pass rush 
We now highlight what these quantities measure as well as their limitations.

[NEED TO ADD "NGS Get Off" https://fivethirtyeight.com/features/the-nfl-has-a-new-way-to-measure-the-explosiveness-of-pass-rushers/]

Perhaps the most commonly-known statistics for evaluating defensive linemen on pass rush plays are sacks, hits, and hurries, which are discretely observed at the play-level.
Officially tracked by the NFL since 1982, a sack is recorded when a defender tackles the QB behind the line of scrimmage before the QB releases a pass.
Sacks represent a simple binary measure of success for a rusher on a pass play.
However, there are many intermediate outcomes for a rusher that can also be considered successes that do not result in a sack.
Other traditional box-score statistics such hits and hurries are collected by various outlets where hits counts the number of times a QB is hit while passing and hurries counts how often a QB has to throw the ball earlier than they would have liked to.
The sum of sacks, hits, and hurries has been defined as "pressures" and is an improvement over sacks, but suffers from problems of subjectivity (e.g. was the play actually a hurry?) and are still binary outcomes (i.e. There is a continuum of plays that aren't hurries that all counted as simply "not hurries").
Pro football focus (PFF) comes up with a metric called "pass-rush productivity" (PRP), which is a minor modification from the aforementioned pressures metric (see Table \ref{tab:metrics}).
In particular, PRP gives twice as much weight to a sack relative to hurries and hits, which is a slight improvement to pressures.
However, the choice of weights is ad-hoc and still only measures binary outcomes similar to the shortcomings of previously discussed metrics.

@@

\begin{table}
\caption{A summary of previously-existing pass rush metrics. \label{tab:metrics}}
\centering
\begin{tabular}{p{0.27\textwidth}p{0.67\textwidth}}
\hline
Metric & Description \\ 
\hline
Sacks & A defender tackles the QB behind the line of scrimmage before the QB is able to throw a pass \\
Hits & A defender tackles the QB behind the line of scrimmage after the QB is able to throw a pass. \\
Hurries & A defender pressures the QB behind the line of scrimmage forcing the QB to throw the ball prior sooner than intended. \\
Pressures & $\text{Hurries} + \text{ Hits } + \text{ Sacks}$ \\
Pass-Rush Productivity & $\displaystyle \frac{(\text{Hurries} + \text{Hits})/2 + \text{Sacks}}{\text{ Pass Rush Snaps}}$\\
Time In Pocket & Time (in seconds) between ball snap and throw or pocket collapse for a QB \\
Pass Rush Win Rate & Rate at which pass rusher beats pass block within 2.5 seconds after ball snap\\
NGS Get Off & add this\\
\hline
\end{tabular}
\end{table}

@@

The peer-reviewed literature on measuring the performance of football linemen (either offensive or defensive) is scant.
@AlamarGould2008 find an association between pass completion rate and successful pass blocking by offensive linemen.
The data for this study are collected for the first three weeks of the 2007 NFL season, manually recording whether a lineman holds a block and the time it took for quarterback to throw the football.
@AlamarGoldner2011 later follow up by using manually-tracked data for the 2010 season to estimate lineman performance for different team-positions instead of individual defenders (e.g. Chicago Bears' left tackle, Pittsburgh Steelers' center).
The paper uses survival analysis to model time in pocket and completion percentage for quarterbacks, before proposing a measure for linemen's contribution to their team's passing in terms of yards gained.

@Wolfson2017Forecasting comment on the two aforementioned articles that "[a]lthough these are exciting preliminary steps, there is still a long way to go before we can provide a comprehensive appraisal of the achievements of an individual lineman."
The challenge here is fundamental, since there were not enough public data at the time to develop any meaningful metric for linemen in football, as also noted by @AlamarGould2008.
However, with the granularity of player tracking data, we have access to data not only for the linemen but also for every player on the field.
This provides us with a great opportunity to assess the performance of linemen in football.

@@

## Our Contribution

In this paper, we focus on the evaluation of defensive linemen in football.
We devise STRAIN, a metric for measuring pass rush effectiveness, inspired by the concept of strain rate in materials science.
Our metric gives a continuous measure of pressure for every pass rusher on the football field over the course of an entire play.
This allows for the assessment of pass rush success even on plays that do not result in an observed outcome like a sack, hit, or hurry.
We view this as a major step forward for accurately evaluating defensive linemen performance.
Further, we note that although our focus in this paper is on defensive linemen, our proposed approach can be extended to the evaluation of offensive linemen in American football.

The remainder of this manuscript is outlined as follows.
We first describe the player tracking data provided by 2023 Big Data Bowl in Section \ref{sec:data}.
We then introduce the mathematical motivation and definition of our statistic STRAIN and our modeling approach in Section \ref{sec:methods}.
Next, we present applications of STRAIN and study different statistical properties of the metric in Section \ref{sec:results}.
We close with our discussion of future directions related to this work in Section \ref{sec:discussion}.
